Co-citations in context

### 2019-06-11 cocit_r.tex
bold restructure
edited last paragraph of introduction

Our analysis of these semantically-themed citation networks challenges the assertion of universality. We demonstrate reduced model misspecification and observe that while HC remains highly correlated with hit articles in the immunology and metabolism datasets, HN is not. \emph{needs another sentence referencing Fig 2}. In addition, while HN is highly correlated with hit articles in applied physics, HC is not. Thus, disciplinary networks are different from each other and the full WoS network. Beyond variation across disciplines and the case for contextual analysis, we found that the categories demonstrating the highest percentage of hits are not robust with respect to varying parameters that define hit articles and the threshold for highly novel citation patterns.

\section{Materials and Methods}
\emph{Bibliographic data} We have previously developed ERNIE, an open source knowledge platform into which we parse the Web of Science (WoS) Core Collection~\citep{Keserci371955}. WoS data stored in ERNIE spans the period 1900-2019 and consists of over 72 million publications. For this study, we generated an analytical dataset from years 1985 to 2005 using data in ERNIE. The total number of publications in this dataset was just over 25 million publications (25,134,073), which were then stratified by year of publication. For each of these years, we further restricted analysis to publications of type Article. Since WoS data also contains incomplete references or references that point at other indexes, we also considered only those references for which there were complete records~(Table \ref{tab:summary_data}). For example, WoS data for year 2005 contained 1,753,174 publications, which after restricting to type Article and considering only those references described above resulted in 916,573 publications, 6,095,594 unique references (set of references), and 17,167,347 total references (multiset of references). Given consistent trends in the data, Table \ref{tab:summary_data}, we analyzed the two boundary years (1985 and 2005) and the mid-point (1995). We also used the number of times each of these articles was cited in the first 8 years since publication as a measure of its impact.

% latex table generated in R 3.5.3 by xtable 1.8-4 package
% Fri May 17 14:28:31 2019
\begin{table}[ht]
\caption{Summary of base WoS Analytical Dataset.The number of publications of type Article (UP), unique references (UR), total references (TR) and the ratio of total references to unique references increases monotonically with each year indicating that both the number of documents and citation activity increase over time. Only publications of type Article with at least two references, and references with complete publication data were included.} 
\label{tab:summary_data}
\centering
\begin{tabular}{|r rrrr|}
  \hline
year & UP & UR & TR & TR/UR \\ 
  \hline
1985 & 391,860 & 2,266,584 & 5,588,861 & 2.47 \\ 
  \hline
1986 & 402,309 & 2,316,451 & 5,708,796 & 2.46 \\ 
1987 & 412,936 & 2,427,347 & 5,998,513 & 2.47 \\ 
1988 & 426,001 & 2,545,647 & 6,354,917 & 2.50 \\ 
1989 & 443,144 & 2,673,092 & 6,749,319 & 2.52 \\ 
1990 & 458,768 & 2,827,517 & 7,209,413 & 2.55 \\ 
1991 & 477,712 & 2,977,784 & 7,729,776 & 2.60 \\ 
1992 & 492,181 & 3,134,109 & 8,188,940 & 2.61 \\ 
1993 & 504,488 & 3,278,102 & 8,676,583 & 2.65 \\ 
1994 & 523,660 & 3,458,072 & 9,255,748 & 2.68 \\ 
  \hline
1995 & 537,160 & 3,680,616 & 9,875,421 & 2.68 \\ 
  \hline
1996 & 663,110 & 4,144,581 & 11,641,286 & 2.81 \\ 
1997 & 677,077 & 4,340,733 & 12,135,104 & 2.80 \\ 
1998 & 693,531 & 4,573,584 & 12,728,629 & 2.78 \\ 
1999 & 709,827 & 4,784,024 & 13,280,828 & 2.78 \\ 
2000 & 721,926 & 5,008,842 & 13,810,746 & 2.76 \\ 
2001 & 727,816 & 5,203,078 & 14,261,189 & 2.74 \\ 
2002 & 747,287 & 5,464,045 & 15,001,390 & 2.75 \\ 
2003 & 786,284 & 5,773,756 & 16,024,652 & 2.78 \\ 
 2004 & 826,834 & 6,095,594 & 17,167,347 & 2.82 \\ 
   \hline
 2005 & 886,648 & 6,615,824 & 19,036,324 & 2.88 \\ 
 \hline
\end{tabular}
\end{table}

We constructed three disciplinary datasets based on keyword searches. (i) immunology (ii) metabolism (iii) applied physics. For the first two, rooted in biomedical research, we searched Pubmed for the term `immunology' or `metabolism' in the years 1985, 1995, and 2005~(Table \ref{tab:disc-datasets}). Pubmed IDs (pmids) returned were matched to WoS IDs (wos\_ids) and used to retrieve relevant articles. For the applied physics dataset, we directly searched traditional subject labels in WoS for `applied physics'. We also examined publications in the five major research areas in the Web of Science; life sciences \& biomedicine, physical sciences, technology, social sciences, arts \& humanities using the extended subcategory classification of 153 sub-groups to categorize disciplinary composition of cited references in the datasets we studied.

\begin{table}[ht]
\caption{Disciplinary Datasets. PubMed and WoS were searched for articles using search terms, `immunology', `metabolism', and `applied physics'. Counts of publications are shown for each of the three years analyzed and expressed in parentheses as a percentage of the total number of publications in our analytical WoS dataset (Table \ref{tab:summary_data}) for that year.}
\label{tab:disc-datasets}
\centering
\begin{tabular}{|r r r r|}
  \hline
Year & Immunology & Metabolism & Applied Physics \\ 
  \hline
1985 & 21,606 (5.5\%) & 78,998 (20.2\%) & 10,298 (2.7\%) \\ 
1995 & 29,320 (5.5\%)  & 121,247 (22.6\%) & 21,012 (3.9\%)  \\ 
2005 & 37,296 (4.2\%) & 200,052 (22.6\%) & 35,600 (4.0\%)  \\ 
 \hline
\end{tabular}
\end{table}

>>>>

Our analyses, using three different techniques, provides substantial evidence that constraining the reference substitutions to the disciplinary subnetwork reduces model mis-specification compared to treating all substitutions within WoS as equiprobable (cite figures). 
Furthermore, re-analyses of the semantically-themed
 citation networks under the improved model reveals strikingly different trends than those 
 observed by Uzzi {\em et al.}. For example, while Uzzi et al. claimed that highly cited articles are more likely to be both HC and HN than expected, and that this trend held across all subdisciplines, we find that these trends vary with the subdiscipline so that no universal trends can be established. Specifically, 
 HC remains highly correlated with highly cited articles in the immunology and metabolism datasets but not with applied physics, and HN is highly correlated with highly cited articles in applied physics but not with immunology and metabolism.    
 Thus, disciplinary networks are different from each other, and trends that hold for the full WoS network do not hold for even large subnetworks (such as metabolism). 
 Furthermore, we also found that the categories HC, HN, etc.,  demonstrating the highest percentage of highly cited articles are not robust with respect to varying thresholds for high citation counts or for highly novel citation patterns.
 Overall, our study, although limited to three subnetworks, suggests that  co-citation analysis, while useful in many regards, cannot be used to provide universal features in scholarship, due to important differences between disciplines and between publications.

\section{Materials and Methods}
\emph{Bibliographic data} We have previously developed ERNIE, an open source knowledge platform into which we parse the Web of Science (WoS) Core Collection~\citep{Keserci371955}. WoS data stored in ERNIE spans the period 1900-2019 and consists of over 72 million publications. For this study, we generated an analytical dataset from years 1985 to 2005 using data in ERNIE. The total number of publications in this dataset was just over 25 million publications (25,134,073), which were then stratified by year of publication. For each of these years, we further restricted analysis to publications of type Article. Since WoS data also contains incomplete references or references that point at other indexes, we also considered only those references for which there were complete records~(Table \ref{tab:summary_data}). For example, WoS data for year 2005 contained 1,753,174 publications, which after restricting to type Article and considering only those references described above resulted in 916,573 publications, 6,095,594 unique references (set of references), and 17,167,347 total references (multiset of references). Given consistent trends in the data, Table \ref{tab:summary_data}, we analyzed the two boundary years (1985 and 2005) and the mid-point (1995). We also used the number of times each of these articles was cited in the first 8 years since publication as a measure of its impact.
\vspace{-4mm}
% latex table generated in R 3.5.3 by xtable 1.8-4 package
% Fri May 17 14:28:31 2019
\begin{table}[ht]
\caption{Summary of base WoS Analytical Dataset.The number of publications of type Article (UP), unique references (UR), total references (TR) and the ratio of total references to unique references increases monotonically with each year indicating that both the number of documents and citation activity increase over time. Only publications of type Article with at least two references, and references with complete publication data were included.} 
\label{tab:summary_data}
\centering
\scalebox{0.7}{
\begin{tabular}{|r rrrr|}
  \hline
year & UP & UR & TR & TR/UR \\ 
  \hline
1985 & 391,860 & 2,266,584 & 5,588,861 & 2.47 \\ 
  \hline
1986 & 402,309 & 2,316,451 & 5,708,796 & 2.46 \\ 
1987 & 412,936 & 2,427,347 & 5,998,513 & 2.47 \\ 
1988 & 426,001 & 2,545,647 & 6,354,917 & 2.50 \\ 
1989 & 443,144 & 2,673,092 & 6,749,319 & 2.52 \\ 
1990 & 458,768 & 2,827,517 & 7,209,413 & 2.55 \\ 
1991 & 477,712 & 2,977,784 & 7,729,776 & 2.60 \\ 
1992 & 492,181 & 3,134,109 & 8,188,940 & 2.61 \\ 
1993 & 504,488 & 3,278,102 & 8,676,583 & 2.65 \\ 
1994 & 523,660 & 3,458,072 & 9,255,748 & 2.68 \\ 
  \hline
1995 & 537,160 & 3,680,616 & 9,875,421 & 2.68 \\ 
  \hline
1996 & 663,110 & 4,144,581 & 11,641,286 & 2.81 \\ 
1997 & 677,077 & 4,340,733 & 12,135,104 & 2.80 \\ 
1998 & 693,531 & 4,573,584 & 12,728,629 & 2.78 \\ 
1999 & 709,827 & 4,784,024 & 13,280,828 & 2.78 \\ 
2000 & 721,926 & 5,008,842 & 13,810,746 & 2.76 \\ 
2001 & 727,816 & 5,203,078 & 14,261,189 & 2.74 \\ 
2002 & 747,287 & 5,464,045 & 15,001,390 & 2.75 \\ 
2003 & 786,284 & 5,773,756 & 16,024,652 & 2.78 \\ 
 2004 & 826,834 & 6,095,594 & 17,167,347 & 2.82 \\ 
   \hline
 2005 & 886,648 & 6,615,824 & 19,036,324 & 2.88 \\ 
 \hline
\end{tabular}}
\vspace{-4mm}%Put here to reduce too much white space after your table 
\end{table}

We constructed three disciplinary datasets based on keyword searches. (i) immunology (ii) metabolism (iii) applied physics. For the first two, rooted in biomedical research, we searched Pubmed for the term `immunology' or `metabolism' in the years 1985, 1995, and 2005~(Table \ref{tab:disc-datasets}). Pubmed IDs (pmids) returned were matched to WoS IDs (wos\_ids) and used to retrieve relevant articles. For the applied physics dataset, we directly searched traditional subject labels in WoS for `applied physics'. We also examined publications in the five major research areas in the Web of Science; life sciences \& biomedicine, physical sciences, technology, social sciences, arts \& humanities using the extended subcategory classification of 153 sub-groups to categorize disciplinary composition of cited references in the datasets we studied.
\vspace{-2 mm}
\begin{table}[ht]
\caption{Disciplinary Datasets. PubMed and WoS were searched for articles using search terms, `immunology', `metabolism', and `applied physics'. Counts of publications are shown for each of the three years analyzed and expressed in parentheses as a percentage of the total number of publications in our analytical WoS dataset (Table \ref{tab:summary_data}) for that year.}
\label{tab:disc-datasets}
\centering
\begin{tabular}{|r r r r|}
  \hline
Year & Immunology & Metabolism & Applied Physics \\ 
  \hline
1985 & 21,606 (5.5\%) & 78,998 (20.2\%) & 10,298 (2.7\%) \\ 
1995 & 29,320 (5.5\%)  & 121,247 (22.6\%) & 21,012 (3.9\%)  \\ 
2005 & 37,296 (4.2\%) & 200,052 (22.6\%) & 35,600 (4.0\%)  \\ 
 \hline
\end{tabular}
\vspace{-6mm}
\end{table}
